---
title: "deepStats - compareCurves"
output: html_notebook
---

/!\ This Notebook is not yet updated /!\

This tool assesses if multiple genomics signals ( ChIP-seq, ATAC-seq... ) are significantly different or
not between conditions ( control, KO1, KO2, etc ). DeepStats compareCurves uses bootstraps and corrected
Wilcoxon Rank-sum tests to do so. The input of this tool corresponds to the output of deepTools

`computeMatrix --outFileNameMatrix`. If multiple region sets have been used in deepTools, one plot and
tab-delimited table will be produced for each set of regions.

## Loading dependencies

Be sure to install all the dependencies. Please check https://github.com/gtrichard/deepStats.

```{r}
suppressMessages( library( 'optparse' ) )
suppressMessages( library( 'boot' ) )
suppressMessages( library( 'ggplot2' ) )
suppressMessages( library( 'cowplot' ) )
suppressMessages( library( 'showtext' ) )
suppressMessages( library( 'tidyverse' ) )
suppressMessages( library( 'dichromat' ) )
suppressMessages( library( 'purrr' ) )
```

## Set working directory

```{r}
setwd("/home/user")
```


## I/O and parameters

```{r}
opt = list()

opt$input = "file.txt"
# DeepTools file obtained from computeMatrix --outFileNameMatrix.
# Alternatively, a .boots file from previous compareCurves runs can be provided
# for replotting purposes and to avoid the bootstraps computation once more.

opt$output = "file"
# Output prefix. Three files will be generated, a .pdf file containing the plot
# and a .boots file containing the bootsraps information ( RDS file ). If a .boots file
# is provided as input, only the plot will be produced as pdf.

opt$comparison = NULL
# When specifying 'regions' or 'scores', force a given comparison. The correct comparison to
# perform is otherwise automatically detected.

opt$scoreLabel = NULL
# Names of the scores.

opt$regionLabel = NULL
# Names of the regions.

opt$bootstraps = 1000
# Number of bootstraps to perform. Default: 1000.

opt$predictionsNumber = NULL
# Number of regions randomly drawn at each bootstrap ( predictions ).
# For more information please see:
# https://www.rdocumentation.org/packages/boot/versions/1.3-20/topics/boot


opt$bootstrapsCI = 0.95
# Confidence intervals ( CI ) threshold in percent. Default: 0.95.

opt$CPU = 4
# Number of CPU to use. Default: 4.

opt$wilcoxStartBin = NULL
# The start bin number used to restrict the corrected
# wilcoxon test to a given portion of the plot. An integer must be provided.
# Default: 1, the first bin.

opt$wilcoxEndBin = NULL
# The end bin number used to restrict the corrected
# wilcoxon test to a given portion of the plot. An integer must be provided.
# Default: the last bin.

opt$wilcoxThreshold = 0.05
# Threshold used to define significant bins on the Wilcoxon rank-sum
# test plot. Default: 0.05

opt$signalName = "Genomic signal"
# Name given to the signal, for instance 'H3K4me3 log2input'.
# Default: 'Genomic signal'

opt$firstRegionName = "TSS"
# Name of the central or left region. Default: 'TSS'.

opt$secondRegionName = "TES"
# Name of the right region, only used when `deeptools computeMatrix`
# ran in scaled-regions mode. Default: 'TES'.

opt$bootPlotColors = NULL
# Change the bootstraps plot color palette to a user-provided one. The file must
# be tab-delimited and contain for each line two HTML color codes ( #3366CC   #769EF2 ).
# The first column corresponds to the mean color, the second column corresponds to the
# color of the bootstrap confidence interval shadowed area. The default color scale
# contains 6 colors that are color blind friendly using the dichromat R package.

opt$bootPlotWidth = 5.2
# How large the bootstraps plot should be. Default: 5.2

opt$bootPlotHeight = 3.7
# How tall the bootstraps plot should be. Default: 3.7

opt$wilcoxPlotWidth = 4.6
# How large the Wilcoxon rank-sum test plot should be. Default: 4.6

opt$wilcoxPlotHeight = 4.6
# How tall the Wilcoxon rank-sum test  plot should be. Default: 4.6

opt$font = NULL
# Font used for plotting, given a TTF file. Default is usually Helvetica.

if ( is.null( opt$input ) ) {
  print_help( opt_parser )
  stop( "An input file must be supplied", call.=FALSE )
}

if ( is.null( opt$output ) ) {
  print_help( opt_parser )
  stop( "An output prefix must be supplied", call.=FALSE )
}

substrRight <- function( x, n ) {
  substr( x, nchar( x )-n+1, nchar( x ) )
}

if ( substrRight( opt$input,6 ) == ".boots"  ) {
  mode="boots"
  cat( as.character( Sys.time() ),".boots file provided as input, skipping statistical analysis.\n" )
}else{mode="deeptools"}

```

## compareCurves analysis

```{r}
##-------------------Set absolute paths-------------------
opt$input <- normalizePath( opt$input )
opt$output <- paste( normalizePath( dirname( opt$output ) ),"/",basename( opt$output ),sep="" )
setwd( normalizePath( dirname( opt$output ) ) )


##-------------------Metadata extraction-------------------
if ( mode == "deeptools" ) {
  master_list <- list()
  metadata <- list()

  cat( as.character( Sys.time() ),"Extracting metadata from file:\n   ", opt$input )

  regions_data <- readLines( opt$input,n = 1 )
  regions_data <-strsplit( x = regions_data,split = "\t" )
  regions_data<-lapply( regions_data,function( x ) strsplit( x,split=":" ) )
  regions_data[[1]][[1]][1] <- gsub( pattern = "#",replacement = "", x = regions_data[[1]][[1]][1] )

  even_indexes<-seq( 2,length( regions_data[[1]] )*2,2 )
  odd_indexes<-seq( 1,length( regions_data[[1]] )*2,2 )

  group_names<-as.character( as.data.frame( unlist( regions_data[[1]] ) )[odd_indexes,] )
  group_length<-as.numeric( as.character( as.data.frame( unlist( regions_data[[1]] ) )[even_indexes,] ) )

  metadata[[1]]<-group_names
  metadata[[2]]<-group_length

  xaxis_data <- readLines( opt$input,n = 2 )[2]
  xaxis_data <-strsplit( x = xaxis_data,split = "\t" )
  xaxis_data <-lapply( xaxis_data,function( x ) strsplit( x,split=":" ) )
  xaxis_data[[1]][[1]][1] <- gsub( pattern = "#",replacement = "", x = xaxis_data[[1]][[1]][1] )

  even_indexes<-seq( 2,length( xaxis_data[[1]] )*2,2 )
  odd_indexes<-seq( 1,length( xaxis_data[[1]] )*2,2 )

  xaxis_names  <- as.character( as.data.frame( unlist( xaxis_data[[1]] ) )[odd_indexes,] )
  xaxis_values <- as.numeric( as.character( as.data.frame( unlist( xaxis_data[[1]] ) )[even_indexes,] ) )

  metadata[[3]]<-xaxis_names
  metadata[[4]]<-xaxis_values

  score_data <- readLines( opt$input,n = 3 )[3]
  score_data <- unlist( strsplit( score_data,'\t' ) )
  score_data<-score_data[-grep( pattern =":",x = score_data )]
  metadata[["scores"]]<-unique( score_data )

  names( metadata )<-c( "group_names","group_length","xaxis_names","xaxis_values","scores" )

  cat( " ... Done!\n" )
}



##-------------------Setup metadata and R objects in case a boots file is provided-------------------
if ( mode == "boots" ) {
  cat( as.character( Sys.time() ),"Extracting data from file:\n   ", opt$input )
  master_list <- readRDS( file=opt$input )
  metadata <- master_list$metadata
  master_list$metadata <- NULL
}


##-------------------Load user-provided labels-------------------
if ( is.null(opt$scoreLabel)==FALSE ) {
  metadata$scores <- unlist(str_split(opt$scoreLabel,";"))
}

if ( is.null(opt$regionLabel)==FALSE ) {
  metadata$group_names <- unlist(str_split(opt$regionLabel,";"))
}


##-------------------Determination of the comparisons to make / load user-provided comparison-------------------
if(is.null(opt$comparison)){

  if( length( metadata$scores )>1 & length( metadata$group_names )==1 ) {
    cat( as.character( Sys.time() ),"Multiple scores detected but only one set of region found: comparing scores.\n" )
    global_flag="compare_scores"
  }

  if( length( metadata$scores )==1 & length( metadata$group_names )>1 ) {
    cat( as.character( Sys.time() ),"Multiple set of regions detected but only one score found: comparing regions.\n" )
    global_flag="compare_regions"
  }

  if( length( metadata$scores )>1 & length( metadata$group_names )>1 ) {
    cat( as.character( Sys.time() ),"Multiple scores and set of regions detected: comparing regions per score.\n" )
    global_flag="compare_scores"
  }

}else{
  if(opt$comparison=="score" || opt$comparison=="scores" || opt$comparison=="Score" || opt$comparison=="Scores"){
    cat( as.character( Sys.time() ),"--comparison is set: comparing scores.\n" )
    global_flag="compare_scores"
  }

  if(opt$comparison=="region" || opt$comparison=="regions" || opt$comparison=="Region" || opt$comparison=="Regions"){
    cat( as.character( Sys.time() ),"--comparison is set: comparing regions.\n" )
    global_flag="compare_regions"
  }
}



##-------------------Bootstrapping-----------------------------
if ( mode=="deeptools" ) {

  bootstraps_to_plot<-list()

  dat <- read.table( file=opt$input, header=F, sep=( "\t" ),skip=3 ) #loading the input file content
  nbbin  <- as.numeric(  ncol( dat  ) / length(  metadata$scores  )  ) # number of bins to analyze ( a bin = a column in the file )
  scores_length <- rep( x=nbbin,times=length( metadata$scores ) )
  line_index_end <- c( cumsum( metadata$group_length ) )
  line_index_start <- c( 1,cumsum( metadata$group_length )+1 )
  col_index_end <- c( cumsum( scores_length ) )
  col_index_start <- c( 1,cumsum( scores_length )+1 )

  cat(  as.character( Sys.time() ), "Boostraps starting with", opt$CPU,"cores, for each group:\n" )
  cat( "   Number of bootstraps to perform =", opt$bootstraps,"\n" )
  cat( "   Number of bins to analyze =", nbbin,"\n" )

  for( score in 1:length( metadata$scores ) ) {

    bootstrap_dfs<-list()

    for ( group in 1:length( metadata$group_names ) ) {

      cat(  as.character( Sys.time() ), "Starting to boostrap values for the following group:\n" )
      cat( "   Genomic score used:",metadata$scores[[score]],"\n" )

      dat_group    <- dat[line_index_start[group]:line_index_end[group],]
      dat_group    <- dat_group[,col_index_start[score]:col_index_end[score]]
      nbgene <- as.numeric( nrow( dat_group ) ) # number of genes/regions in the dataset
      label  <- metadata$group_names[[group]]
      output.name <- opt$output

      cat( "   Region set used:",label,"\n" )
      cat( "   Number of regions to analyze:", nbgene )

      sample.colmeans <- function( x, d ) {
        return( apply( x[d,], 2, mean, na.rm=T ) )
      }

      b = boot( dat_group, sample.colmeans, R=opt$bootstraps, ncpus=opt$CPU, parallel = "multicore" )

      mini <- vector( length = nbbin )
      maxi <- vector( length = nbbin )

      ci_min <- ( 1-opt$bootstrapsCI )/2
      ci_max <- opt$bootstrapsCI+ci_min

      for ( i in 1:nbbin ) {
        mini[i] <- quantile( b$t[,i], ci_min, na.rm=T )           # get CI
        maxi[i] <- quantile( b$t[,i], ci_max, na.rm=T )
      }

      cat( "... Done!\n" )

      region_name<-rep( x=label,nbbin )
      score_name<-rep( x=metadata$scores[[score]],nbbin )
      bin<-seq( from=1, to=nbbin, by=1 )


      bootstraps_to_plot<-cbind( region_name,score_name,bin,b$t0,mini,maxi )
      bootstraps_to_plot<-as.data.frame( unname( bootstraps_to_plot ) )
      colnames( bootstraps_to_plot )<-c("Region","Score","Bin","Mean","Min","Max" )

      for ( column in 3:ncol( bootstraps_to_plot ) ) {
        bootstraps_to_plot[,column]<-as.numeric( as.character( bootstraps_to_plot[,column] ) )
      }

      bootstrap_dfs[[group]] <- bootstraps_to_plot
    }
    names( bootstrap_dfs )<- metadata$group_names

    #bootstraps_to_plot <- do.call( rbind,bootstrap_dfs )

    master_list[[metadata$scores[score]]] <- list()
    master_list[[score]][["bootstraps"]] <- bootstrap_dfs
  }
}


##-------------------Wilcoxon rank-sum test-------------------

global_flag="compare_regions"

if ( mode=="deeptools" ) {

  if ( is.null( opt$wilcoxStartBin )==TRUE && is.null( opt$wilcoxEndBin )==FALSE ) {
    print_help( opt_parser )
    stop( "Both --wilcoxStartBin and --wilcoxEndBin must be given.", call.=FALSE )
  }

  if ( is.null( opt$wilcoxStartBin )==FALSE && is.null( opt$wilcoxEndBin )==TRUE ) {
    print_help( opt_parser )
    stop( "Both --wilcoxStartBin and --wilcoxEndBin must be given.", call.=FALSE )
  }

  if ( is.null( opt$wilcoxStartBin )==FALSE && is.null( opt$wilcoxEndBin )==FALSE ) {
    if ( opt$wilcoxStartBin > opt$wilcoxEndBin ) {
      print_help( opt_parser )
      stop( "--wilcoxStartBin must be superior to --wilcoxEndBin.", call.=FALSE )
    }
    if ( opt$wilcoxStartBin > opt$wilcoxEndBin ) {
      cat( as.character( Sys.time() ), "Starting Wilcoxon rank-sum test on the following bins: from",opt$wilcoxStartBin,
           "to",opt$wilcoxEndBin, "\n" )
      wilcox_df <- dat[,opt$wilcoxStartBin:opt$wilcoxEndBin]
      #wilcox_dfs <- lapply( bootstrap_dfs, function( x ) {x[which( x$Bin >= opt$wilcoxStartBin & x$Bin <= opt$wilcoxEndBin  ),]} )
    }
  }else{
    cat( as.character( Sys.time() ), "Starting Wilcoxon rank-sum test.\n" )
    wilcox_df <- dat
  }

  if (global_flag=="compare_regions") {

    for( score in 1:length( metadata$scores ) ) {

      wilcox_groups_indexes <- paste( line_index_start[1:length( line_index_start )-1],line_index_end,sep=":" )
      wilcox_comparisons <- as.data.frame( expand.grid( wilcox_groups_indexes,wilcox_groups_indexes ) )

      wilcox_score_indexes <- paste( col_index_start[1:length( col_index_start )-1],col_index_end,sep=":" ) #this is type dependent

      wilcox_matrix <- matrix( nrow = length( wilcox_groups_indexes ), ncol = length( wilcox_groups_indexes ) )
      row.names( wilcox_matrix )<-wilcox_groups_indexes
      colnames( wilcox_matrix )<-wilcox_groups_indexes

      wilcox_padj_vectors <- list()

      for ( comps in ( 1:nrow( wilcox_comparisons ) ) ) {

        group_1_name  <- as.character( wilcox_comparisons[comps,1] )
        group_2_name  <- as.character( wilcox_comparisons[comps,2] )

        group_1_index <- unlist( strsplit( group_1_name,split = ":" ) )
        group_2_index <- unlist( strsplit( group_2_name,split = ":" ) )

        score_index <- unlist( strsplit( as.character(wilcox_score_indexes[score]),split = ":" ) )

        group_1_data  <- wilcox_df[group_1_index[1]:group_1_index[2],score_index[1]:score_index[2]]
        group_2_data  <- wilcox_df[group_2_index[1]:group_2_index[2],score_index[1]:score_index[2]]

        wilcox_res <- vector()

        for ( bins in ( 1:ncol( group_1_data ) ) ) {
          wilcox_res[bins] <- wilcox.test( group_1_data[,bins],group_2_data[,bins] )$p.value
        }

        wilcox_adjust <- p.adjust( wilcox_res,method = "bonferroni" )
        wilcox_padj_vectors[[ paste( group_1_name,group_2_name,sep="_" )]] <- wilcox_adjust
      }

      max_val=max( as.vector( unlist( lapply( wilcox_padj_vectors, function( x ) max( -log10( x ) ) ) ) ) )

      wilcox_plot_list <- list()
      wilcox_to_plot <- list()

      group_names_mapping <- as.data.frame( unique( wilcox_comparisons[,1] ) )

      group_names_mapping["names"]   <- metadata$group_names
      colnames( group_names_mapping )<-c( "indexes","names" )

      for ( i in 1:length( wilcox_padj_vectors ) ) {
        wilcox_to_plot[[i]]<-as.data.frame( -log10( wilcox_padj_vectors[[i]] ) )
        wilcox_to_plot[[i]]["xaxis"]<-as.numeric( as.character( rownames( wilcox_to_plot[[i]] ) ) )
        colnames( wilcox_to_plot[[i]] )<-c( "vals","xaxis" )

        wilcox_plot_group1<-as.data.frame( unlist( strsplit( names( wilcox_padj_vectors )[i],split = "_" ) )[1] )
        colnames( wilcox_plot_group1 )<-"indexes"
        wilcox_plot_group1<- merge( wilcox_plot_group1,group_names_mapping,by="indexes" )[1,2]

        wilcox_plot_group2<-as.data.frame( unlist( strsplit( names( wilcox_padj_vectors )[i],split = "_" ) )[2] )
        colnames( wilcox_plot_group2 )<-"indexes"
        wilcox_plot_group2<- merge( wilcox_plot_group2,group_names_mapping,by="indexes" )[1,2]

        wilcox_to_plot[[i]]["Group1"]<-wilcox_plot_group1
        wilcox_to_plot[[i]]["Group2"]<-wilcox_plot_group2
        wilcox_to_plot[[i]]["Constant"]<-metadata$scores[[score]]
      }
      master_list[[score]][["wilcoxon"]] <- wilcox_to_plot
    }
  }

  if (global_flag=="compare_scores") {

    store_wilcox_scores<-list()

    for( region in 1:length( metadata$group_names ) ) {

      wilcox_score_indexes <- paste( col_index_start[1:length( col_index_start )-1],col_index_end,sep=":" ) #this is type dependent
      wilcox_comparisons <- as.data.frame( expand.grid( wilcox_score_indexes,wilcox_score_indexes ) ) #this is type dependent

      wilcox_groups_indexes <- paste( line_index_start[1:length( line_index_start )-1],line_index_end,sep=":" )

      wilcox_matrix <- matrix( nrow = length( wilcox_score_indexes ), ncol = length( wilcox_score_indexes ) )
      row.names( wilcox_matrix )<-wilcox_score_indexes
      colnames( wilcox_matrix )<-wilcox_score_indexes

      wilcox_padj_vectors <- list()

      for ( comps in ( 1:nrow( wilcox_comparisons ) ) ) {

        group_1_name  <- as.character( wilcox_comparisons[comps,1] )
        group_2_name  <- as.character( wilcox_comparisons[comps,2] )

        group_1_index <- unlist( strsplit( group_1_name,split = ":" ) )
        group_2_index <- unlist( strsplit( group_2_name,split = ":" ) )

        regions_index <- unlist( strsplit( wilcox_groups_indexes[region],split = ":" ) )

        group_1_data  <- wilcox_df[regions_index[1]:regions_index[2],group_1_index[1]:group_1_index[2]] #this is type dependent
        group_2_data  <- wilcox_df[regions_index[1]:regions_index[2],group_2_index[1]:group_2_index[2]] #this is type dependent

        wilcox_res <- vector()

        for ( bins in ( 1:ncol( group_1_data ) ) ) {
          wilcox_res[bins] <- wilcox.test( group_1_data[,bins],group_2_data[,bins] )$p.value
        }

        wilcox_adjust <- p.adjust( wilcox_res,method = "bonferroni" )
        wilcox_padj_vectors[[ paste( group_1_name,group_2_name,sep="_" )]] <- wilcox_adjust
      }

      max_val=max( as.vector( unlist( lapply( wilcox_padj_vectors, function( x ) max( -log10( x ) ) ) ) ) )

      wilcox_plot_list <- list()
      wilcox_to_plot <- list()

      group_names_mapping <- as.data.frame( unique( wilcox_comparisons[,1] ) )

      group_names_mapping["names"]   <- metadata$scores #this is type dependent
      colnames( group_names_mapping )<-c( "indexes","names" ) #this is type dependent

      for ( i in 1:length( wilcox_padj_vectors ) ) {
        wilcox_to_plot[[i]]<-as.data.frame( -log10( wilcox_padj_vectors[[i]] ) )
        wilcox_to_plot[[i]]["xaxis"]<-as.numeric( as.character( rownames( wilcox_to_plot[[i]] ) ) )
        colnames( wilcox_to_plot[[i]] )<-c( "vals","xaxis" )

        wilcox_plot_group1<-as.data.frame( unlist( strsplit( names( wilcox_padj_vectors )[i],split = "_" ) )[1] )
        colnames( wilcox_plot_group1 )<-"indexes"
        wilcox_plot_group1<- merge( wilcox_plot_group1,group_names_mapping,by="indexes" )[1,2]

        wilcox_plot_group2<-as.data.frame( unlist( strsplit( names( wilcox_padj_vectors )[i],split = "_" ) )[2] )
        colnames( wilcox_plot_group2 )<-"indexes"
        wilcox_plot_group2<- merge( wilcox_plot_group2,group_names_mapping,by="indexes" )[1,2]

        wilcox_to_plot[[i]]["Group1"]<-wilcox_plot_group1
        wilcox_to_plot[[i]]["Group2"]<-wilcox_plot_group2
        wilcox_to_plot[[i]]["Constant"]<-metadata$group_names[[region]]
      }
      store_wilcox_scores[[region]]<- do.call( "rbind",wilcox_to_plot )
    }
    for( score in 1:length( metadata$scores ) ) {
      master_list[[score]][["wilcoxon"]] <- store_wilcox_scores
    }
  }
  cat( as.character( Sys.time() ), "Wilcoxon test finished.\n" )
}


```

## Exportation of the data for future runs

```{r}
if ( mode=="deeptools" ) {
  cat( as.character( Sys.time() ),"Exporting data to the following RDS object:\n   ",paste( opt$output,".boots",sep="" ) )
  to_export <- master_list
  to_export[["metadata"]]<-metadata
  saveRDS( to_export,file = paste( opt$output,".boots",sep="" ) )
  rm(to_export)
  cat( "  ... Done!\n" )
}
```

## X axis creation for plotting

```{r}
cat( as.character( Sys.time() )," Generating the X axis for plots.",sep="" )

xaxis<-list()

linebreaker <- function( x,... ) {
  gsub( '\\s','\n',x )
}

if ( metadata$xaxis_values[3]==0 ) {
  #Centered region
  #Start and end definition
  xaxis["xaxis_start"][1] <- as.numeric( 1 ) # position
  xaxis$xaxis_start[2] <- paste( ( -1 * metadata$xaxis_values[2] / 1000 ),"kb",sep = "" )
  xaxis["xaxis_end"][1] <- as.numeric( max( bootstraps_to_plot$Bin ) ) # position
  xaxis$xaxis_end[2] <- paste( ( metadata$xaxis_values[1] / 1000 ),"kb",sep = "" )
  xaxis["first_region"][1] <- round( as.integer( max( bootstraps_to_plot$Bin ) ) / 2 )
  xaxis$first_region[2] <- linebreaker( opt$firstRegionName )
  xaxis$second_region <- NULL

}else{
  #Scaled region
  xaxis["xaxis_start"][1] <- 0 # position
  xaxis$xaxis_start[2] <- paste( ( -1 * metadata$xaxis_values[2] / 1000 ),"kb",sep = "" )
  xaxis["xaxis_end"][1] <- as.numeric( max( bootstraps_to_plot$Bin ) ) # position
  xaxis$xaxis_end[2] <- paste( ( metadata$xaxis_values[1] / 1000 ),"kb",sep = "" )
  xaxis["first_region"] <- metadata$xaxis_values[2] /  metadata$xaxis_values[4]
  xaxis$first_region[2] <- linebreaker( opt$firstRegionName )
  xaxis["second_region"] <- as.numeric(xaxis$xaxis_end[1]) - (metadata$xaxis_values[1]) /  metadata$xaxis_values[4]
  xaxis$second_region[2] <- linebreaker( opt$secondRegionName )
}

if ( metadata$xaxis_values[5]>0 ) {
  xaxis["unscaled5"][1] <- (metadata$xaxis_values[2] + metadata$xaxis_values[5]) / metadata$xaxis_values[4]
  xaxis$unscaled5[2] <- paste( ( -1 * metadata$xaxis_values[5] / 1000 ),"kb",sep = "" )
}

if ( metadata$xaxis_values[6]>0 ) {
  xaxis["unscaled3"][1] <- as.numeric(xaxis$xaxis_end[1]) - (metadata$xaxis_values[1] + metadata$xaxis_values[6]) / metadata$xaxis_values[4]
  xaxis$unscaled3[2] <- paste( ( -1 * metadata$xaxis_values[6] / 1000 ),"kb",sep = "" )
}

xaxis["breaks"]<-as.data.frame( as.numeric( c(  xaxis$xaxis_start[1],
                                                xaxis$unscaled5[1],
                                                xaxis$first_region[1],
                                                xaxis$second_region[1],
                                                xaxis$unscaled3[1],
                                                xaxis$xaxis_end[1] ) ) )

xaxis["names"]<-as.data.frame( as.character( c(  xaxis$xaxis_start[2],
                                                 xaxis$unscaled5[2],
                                                 xaxis$first_region[2],
                                                 xaxis$second_region[2],
                                                 xaxis$unscaled3[2],
                                                 xaxis$xaxis_end[2] ) ) )

cat( "  ... Done!\n" )

```

## Bootstraps plotting

```{r}
cat( as.character( Sys.time() ), "Starting Bootstraps plotting to the following file:\n   ",paste( opt$output,"_bootstraps.pdf",sep="" ) )

pdf( file = paste( opt$output,"_bootstraps.pdf",sep="" ),width = opt$bootPlotWidth,height = opt$bootPlotHeight )

color_palette<-c( "#000000",
                  colorschemes$Categorical.12[[12]],
                  colorschemes$Categorical.12[[8]],
                  colorschemes$Categorical.12[[2]],
                  colorschemes$Categorical.12[[10]],
                  colorschemes$Categorical.12[[6]] )

fill_palette<-c( "#919191",
                 colorschemes$Categorical.12[[11]],
                 colorschemes$Categorical.12[[7]],
                 colorschemes$Categorical.12[[1]],
                 colorschemes$Categorical.12[[9]],
                 colorschemes$Categorical.12[[5]] )

boot_plot_data <- list()

for (group in 1:length(metadata$group_names)){ # One plot per region
  data_frame_storage <- list()


  for (score in 1:length(metadata$scores)) {
    score_name <- metadata$scores[score]
    group_name <- metadata$group_names[group]
    data_frame_storage[[score]] <- master_list[[score_name]]$bootstraps[[group_name]]
  }

  bootstraps_to_plot <- do.call( rbind,data_frame_storage)
  boot_plot_data[[group]]<-bootstraps_to_plot
}

bootstraps_to_plot<-do.call("rbind", boot_plot_data)
rm(boot_plot_data)

make_boot_plot<-function(x,type){
  ggplot(x,aes( x=x$Bin, color=x[,type], fill=x[,type] ))+
    geom_ribbon( aes( ymin = x$Min, ymax = x$Max ),alpha=0.5,linetype=0 )+
    geom_line( aes( y=x$Mean ) )+
    scale_color_manual( "", values = color_palette )+
    scale_fill_manual( "", values = fill_palette )+
    scale_x_continuous( NULL , breaks=xaxis$breaks, labels=xaxis$names )+
    ylab( opt$signalName )+
    panel_border( size=1, remove=FALSE )+
    theme( axis.text.y = element_text( size = 12, color = "#383838" ),
           axis.text.x = element_text( size = 12,  color = "#383838" ),
           axis.title.x = element_text( size = 13 ),
           plot.title = element_text( size = 13, face = "plain" ),
           line = element_line( size=1 ),
           strip.placement = "outside",
           strip.background = element_blank(),
           plot.background = element_rect( colour = "black" ),
           panel.border = element_rect( colour = "black" ),
           axis.line = element_blank(),
           panel.spacing = unit( 1, "line" ) )
}

boot_plot_list <- list()

if(global_flag == "compare_regions") {
  for (score in 1:length(unique(bootstraps_to_plot$Score))){
    data <- bootstraps_to_plot[which(bootstraps_to_plot$Score==unique(bootstraps_to_plot$Score)[score]),]
    boot_plot_list[[score]] <- make_boot_plot(data,"Region") + ggtitle(as.character(unique(data$Score)))
    if ( is.null( opt$font )==FALSE ) {
      font_add( family = "plot_font", regular = opt$font )
      showtext_auto()
      boot_plot_list[[score]] <- boot_plot_list[[score]] + theme( text = element_text( family = "plot_font" ) )
    }
    if( is.null( opt$bootPlotColors )==FALSE ) {
      color_pal<-read.table( opt$bootPlotColors,header = F,sep = "\t" )
      boot_plot_list[[score]] <- boot_plot_list[[score]] +
        scale_color_manual( "",values = color_pal[,1] ) +
        scale_fill_manual( "",values = color_pal[,2] )
    }
  }
}


if(global_flag == "compare_scores") {
  for (region in 1:length(unique(bootstraps_to_plot$Region))){
    data <- bootstraps_to_plot[which(bootstraps_to_plot$Region==unique(bootstraps_to_plot$Region)[region]),]
    boot_plot_list[[region]] <- make_boot_plot(data,"Score") + ggtitle(as.character(unique(data$Region)))
    if ( is.null( opt$font )==FALSE ) {
      font_add( family = "plot_font", regular = opt$font )
      showtext_auto()
      boot_plot_list[[score]] <- boot_plot_list[[score]] + theme( text = element_text( family = "plot_font" ) )
    }
    if( is.null( opt$bootPlotColors )==FALSE ) {
      color_pal<-read.table( opt$bootPlotColors,header = F,sep = "\t" )
      boot_plot_list[[score]] <- boot_plot_list[[score]] +
        scale_color_manual( "",values = color_pal[,1] ) +
        scale_fill_manual( "",values = color_pal[,2] )
    }
  }
}


invisible( boot_plot_list )

invisible( dev.off() )

cat( "  ... Done!\n" )

```

## Wilcoxon rank-sum test plotting

```{r}
cat( as.character( Sys.time() ), "Starting Wilcoxon corrected tests plotting to the following file:\n   ",paste( opt$output,"_wilcoxon.pdf",sep="" ) )

rec_start<-as.numeric( xaxis$xaxis_start[[1]] )
rec_end<-as.numeric( xaxis$xaxis_end[[1]] )

pdf( file = paste( opt$output,"_wilcoxon.pdf",sep="" ),width = opt$wilcoxPlotWidth,height = opt$wilcoxPlotHeight )

make_wilcox_plot<-function(x){
  ggplot( x,aes( x=x$xaxis,y=x$vals ) )+
    geom_rect( aes( xmin = rec_start, xmax = rec_end, ymin = -0.1, ymax =-log10( opt$wilcoxThreshold ) ), fill = "gray" ) +
    geom_line( color="#ff6100" )+
    facet_grid( Group1 ~ Group2 )+
    scale_x_continuous( "Genomic position",breaks=xaxis$breaks,labels=xaxis$names )+
    ylab( "-log10( Bonferroni corrected \n Wilcoxon rank-sum p-value )" )+
    panel_border( size=1,remove=FALSE )+
    theme( axis.text = element_text( size=11 ),
           axis.text.y = element_text( size=10 ),
           axis.text.x = element_text( size=10,angle=50,hjust=1,vjust=1 ),
           plot.title = element_text( size = 13, face = "plain" ),
           line = element_line( size=1 ),
           strip.placement = "outside",
           strip.background = element_blank(),
           plot.background = element_rect( colour = "black" ),
           panel.border = element_rect( colour = "black" ),
           axis.line = element_blank(),
           panel.spacing = unit( 1,"line" ) )

}

wilcox_plot_list <- list()

if(global_flag == "compare_scores") {
  wilcox_to_plot <- master_list[[1]]$wilcoxon
  for (i in 1:length(wilcox_to_plot)) {
    wilcox_plot_list[[i]] <- make_wilcox_plot(wilcox_to_plot[[i]]) + ggtitle(unique(wilcox_to_plot[[i]]$Constant))
    if ( is.null( opt$font )==FALSE ) {
      font_add( family = "plot_font", regular = opt$font )
      showtext_auto()
      wilcox_plot_list[[i]] <- wilcox_plot_list[[i]] + theme( text = element_text( family = "plot_font" ) )
    }
  }
}

if(global_flag == "compare_regions") {
  for (score in 1:length(master_list)){
    wilcox_to_plot<-do.call(rbind,master_list[[score]]$wilcoxon)
    wilcox_plot_list[[score]] <- make_wilcox_plot(wilcox_to_plot) + ggtitle(unique(wilcox_to_plot$Constant))
    if ( is.null( opt$font )==FALSE ) {
      font_add( family = "plot_font", regular = opt$font )
      showtext_auto()
      wilcox_plot_list[[score]] <- wilcox_plot_list[[score]] + theme( text = element_text( family = "plot_font" ) )
    }
  }
}

invisible( wilcox_plot_list )

invisible( dev.off() )

cat( "  ... Done!\n" )
```
